---
title: "Survival Analysis of Cancer Patients"
author: "Akshay"
date: "2026-03-26"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

## Introduction

This project analyzes survival patterns of cancer patients using Kaplan-Meier estimation and Cox proportional hazards regression. The objective is to estimate survival probabilities and identify factors affecting patient survival.

## Load Required Libraries

```{r}
library(survival)
library(survminer)

```

## Load Dataset

```{r}
data <- read.csv(file.choose())
head(data)
names(data)
nrow(data)
```

## Data Overview

```{r}
summary(data)
table(data$Status)

```

## Kaplan-Meier Survival Analysis (Overall)

```{r}
fit <- survfit(Surv(Survival_Time_Days, Status) ~ 1, data = data)
summary(fit)

```

## Plot Kaplan-Meier Survival Curve

```{r}
ggsurvplot(fit, data = data, xlab = "Time (Days)", ylab = "Survival Probability",title = "Kaplan-Meier Survival Curve",conf.int = TRUE)

```

## Kaplan-Meier Survival by Cancer Stage

```{r}
fit_stage <- survfit(Surv(Survival_Time_Days, Status) ~ Cancer_Stage, data = data)

ggsurvplot(fit_stage, data = data, pval = TRUE, legend.title = "Cancer Stage", xlab = "Time (Days)", ylab = "Survival Probability")


```

## Cox Proportional Hazards Model

```{r}
cox_model <- coxph(Surv(Survival_Time_Days, Status) ~ Cancer_Stage + Age + Gender, data = data)
summary(cox_model)

```

## Life Table (Interval Survival Estimates)

```{r}
summary(fit, times = seq(0, max(data$Survival_Time_Days), by = 100))

```

## Conclusion

The Kaplan-Meier analysis shows a steady decline in survival probability over time. The median survival is approximately 1330 days. Cox regression indicates that cancer stage is a significant predictor of survival, highlighting the importance of early diagnosis.
